Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2017-02-08 manuscript/editor_response_tree_physiology.Rmd
editor response finished and sent to Remko

We are grateful for the opportunity to resubmit our manuscript. We thank the editor for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from the editor.

- *Comment: 1. It seems that there are two different effects: rooting volume and container/free rooting. And it seems that the latter is dominating and the clear relations with rooting volume are often missing... ignoring this later treatment may depply change your way of interpreting the results. Even the stats are strongly influenced by this treatment (effect on residual error for example).*

\
*Response*: We agree with the editor's comments that the effects of soil volume have not been properly accounted for in several places in the manuscript. Particularly in the discussion we have clarified the nature of the effect type, where soil volume effects have often been over-emphasized. We agree that the stats are strongly influenced by the Free seedling, but this is also representative of the observed effects of root restriction on *A*~sat~, TNC and leaf N across all containers throughout all measurement campaigns. As the impacts of root restriction and poor quality soil impacted were evident across all treatments 
We feel that addressing individual effects of specific volume treatments, or combining soil volume treatment effects, would overly complicate the paper and would not improve our interpretation of the results.

@@

\
*Response*: The editor makes several good points here. We have re-written the equation to better describe the way daily total leaf net C gain (*P*~day~) was calculated. We have also removed the statement "Leaf respiration is included in the calculation of *P*." and added a accurate description of how the area based respiration component was calculated at each 15 minute interval to then model leaf level net photosynthesis rates.

\
- *Comment: 7. L264: How do simulated An values fit with measured Asat?*

\
*Response*:

@@

\
*Response*: We agree with each of the editors points here.  With a comment #6 we have now clearly defined the "total net leaf C gain term".  As the editor points out there are several places where leaf *A*~n~ is discussed inappropriately, especially in the discussion.  Throughout the manuscript we have revised use of terms for C gain to accurately reflect either measured photosynthesis rates (*A*~sat~), leaf level rates of net photosynthesis (*A*~n~) or modelled cumulative total net leaf C gain (*P*~day~).

@@

\
- *Comment: 11. L349-350: These sentences are confusing: does the first sentence refer to Amax on a leaf area basis. Please clarify.*

@@

\
- *Comment: 12. L351: I will not say that Amass is highly correlated with both leaf Nf and leaf starch, when it only accounts for 28% and 11% of the total variations.*

@@

\
- *Comment: 13. L359: I am not convinced that Amax was positively correlated with final harvest biomass. The correlation is driven by the free rooting values. Variation in Amax across rooting volume does not reflect those of biomass in the last point is removed. It means that the co-variation is not so much driven by soil volume. The only clear effect is between container/ free rooting. But it is no more a correlation.*

@@

\
- *Comment: 15. L394: Leaf An (an leaf area basis) or whole plant C gain? Please clarify.*

\
*Response*: editorial change made.  Change *A*~n~ to *A*~sat (measured values)

@@

\
- *Comment: 17. L410: this is not true. There is large reduction in biomass related to container size, but not so much effect of container size on Amax.*

\
*Response*: We have revised this statement to make the point that reductions in measured *A*~sat~ and seedling allometry occurred simultaneosly during early measurement campaigns, and these reductions where sustained throughout the experiement.

\
- *Comment: 18. L413: what is An here? Leaf area or whole plant based?*

\
*Response*: editorial change made.  Changed *A*~n~ to *A*~sat (measured values).

\
- *Comment: 19. L430-431. Correlation between leaf N and An (what is it?) has not been shown in the result sections.*

@@

\
*Response*: We agree with the editor. We have changed this sentence to give a quantitative description of root N content differences between roots of Free seedlings and seedlings in containers (+15 %). 
\
- *Comment: 21. L441: You did not measure available soil N pool but total soil N, which is different. It would have been interesting to know if microbial N mineralization is affected by the containers or not. The key might not be at the plant level but at the microbe level. And what about other nutrients such as available P?*

@@

\
- *Comment: 22. L443: again, what is An? And please show us the evidence of a sink limitation on An of seedlings in container. There is not so much effect of container size on leaf photosynthetic parameter in Table 3*

\
*Response*: editorial change made.  Changed *A*~n~ to *A*~sat (measured values). As with comment #10, *A*~sat values are added in Table 3 to show reductions in measured leaf level photosynthesis in containerized seedlings.

\
- *Comment: 23. L449: I am not convinced that the decline in Amax is correlated with higher starch content, at least not across the different container size. This is not really an effect of rooting volume but again container versus free soil.*

@@

\
- *Comment: 25. L483: There is no reduction in leaf photosynthesis due to the restricted soil volume, according to the data you reported in Table 3. Make clear what is leaf An. The reduction in whole plant assimilation may mainly be due to a reduction in plant leaf area, which in fact is only evident for the largest soil volume. Final plant leaf area can be added in table 1 with stats. It will be helpful as it is not so easy to visualize differences on figure 1.*

@@

\
- *Comment: 26. L477: L483-484: I disagree that reductions in Amax and biomass were strongly correlated. The correlation is driven by high value of the free soil. R2 will drop from 0.88 to 0.32 if you remove this last treatment.*

>>>>

\
\
We are grateful for the opportunity to resubmit our manuscript. We thank the editor for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from the editor.

\
- *Comment: 1. It seems that there are two different effects: rooting volume and container/free rooting. And it seems that the latter is dominating and the clear relations with rooting volume are often missing... ignoring this later treatment may deeply change your way of interpreting the results. Even the stats are strongly influenced by this treatment (effect on residual error for example).*

\
*Response*: We agree with the editor's comments that the effects of soil volume have not been properly accounted for in several places in the manuscript. Particularly in the discussion we have clarified the nature of the effect type, where soil volume effects were often been over-emphasized. We agree that the stats are strongly influenced by the Free seedling, but this is also representative of the observed effects of root restriction on *A*~sat~, TNC and leaf N across all containers throughout all measurement campaigns. We feel that addressing individual effects of specific volume treatments would overly complicate the paper and combining all soil volume treatment effects would not improve our interpretation of the results.

@@

\
*Response*: The editor makes several good points here. We have re-written the equation to better describe the way daily total leaf net C gain (*P*~day~) was calculated. We have also removed the statement "Leaf respiration is included in the calculation of *P*." and added a accurate description of how the area based respiration component was calculated at each 15 minute interval in the modelling of leaf net photosynthesis rates.

\
- *Comment: 7. L264: How do simulated An values fit with measured *A*~sat~?*

\
*Response*: It is difficult to accurately compare spot measurements of *A*~sat~ at saturating light inside the gas exchange cuvette with modelled values of *A*~n~ that are estimated over continually fluctuating meteorological conditions over 120 days. With simple modelling exercise, using Licor conditions and measured photosynthetic parameters the estimated rates are slightly higher than observed rates of *A*~sat~. Importantly, this does not effect the overall patterns of decreased CUE in containerized seedlings.

@@

\
*Response*: We agree with each of the editors points here.  With comment #6 we have now clearly defined the "total net leaf C gain term".  As the editor points out there are several places where leaf *A*~n~ is discussed inappropriately, especially in the discussion.  Throughout the manuscript we have revised use of terms for C gain to accurately reflect either measured photosynthesis rates (*A*~sat~), leaf level rates of net photosynthesis (*A*~n~) or modelled cumulative total net leaf C gain (*P*~day~).

@@

\
- *Comment: 11. L349-350: These sentences are confusing: does the first sentence refer to *A*~max~ on a leaf area basis. Please clarify.*

@@

\
- *Comment: 12. L351: I will not say that A~mass~ is highly correlated with both leaf N~f~ and leaf starch, when it only accounts for 28% and 11% of the total variations.*

@@

\
- *Comment: 13. L359: I am not convinced that Amax was positively correlated with final harvest biomass. The correlation is driven by the free rooting values. Variation in A~mass~ across rooting volume does not reflect those of biomass in the last point is removed. It means that the co-variation is not so much driven by soil volume. The only clear effect is between container/ free rooting. But it is no more a correlation.*

@@

\
- *Comment: 15. L394: Leaf A~n~ (an leaf area basis) or whole plant C gain? Please clarify.*

\
*Response*: editorial change made.  Change *A*~n~ to *A*~sat~ (measured values)

@@

\
- *Comment: 17. L410: this is not true. There is large reduction in biomass related to container size, but not so much effect of container size on A~max~.*

\
*Response*: We have revised this statement to make the point that reductions in measured *A*~sat~ and seedling allometry occurred simultaneously during early measurement campaigns, and these reductions where sustained throughout the experiment.

\
- *Comment: 18. L413: what is A~n~ here? Leaf area or whole plant based?*

\
*Response*: editorial change made.  Changed *A*~n~ to *A*~sat~ (measured values).

\
- *Comment: 19. L430-431. Correlation between leaf N and A~n~ (what is it?) has not been shown in the result sections.*

@@

\
*Response*: We agree with the editor. We have changed this sentence to give a quantitative description of root N content differences between roots of Free seedlings and seedlings in containers (+15 %).

\
- *Comment: 21. L441: You did not measure available soil N pool but total soil N, which is different. It would have been interesting to know if microbial N mineralization is affected by the containers or not. The key might not be at the plant level but at the microbe level. And what about other nutrients such as available P?*

@@

\
- *Comment: 22. L443: again, what is A~n~? And please show us the evidence of a sink limitation on A~n~ of seedlings in container. There is not so much effect of container size on leaf photosynthetic parameter in Table 3*

\
*Response*: editorial change made.  Changed *A*~n~ to *A*~sat~ (measured values). As with comment #10, *A*~sat~ values are also added in Table 3 to show reductions in measured leaf level photosynthesis in containerized seedlings.

\
- *Comment: 23. L449: I am not convinced that the decline in A~max~ is correlated with higher starch content, at least not across the different container size. This is not really an effect of rooting volume but again container versus free soil.*

@@

\
- *Comment: 25. L483: There is no reduction in leaf photosynthesis due to the restricted soil volume, according to the data you reported in Table 3. Make clear what is leaf A~n~. The reduction in whole plant assimilation may mainly be due to a reduction in plant leaf area, which in fact is only evident for the largest soil volume. Final plant leaf area can be added in table 1 with stats. It will be helpful as it is not so easy to visualize differences on figure 1.*

@@

\
- *Comment: 26. L477: L483-484: I disagree that reductions in A~max~ and biomass were strongly correlated. The correlation is driven by high value of the free soil. R2 will drop from 0.88 to 0.32 if you remove this last treatment.*

### 2017-02-08 manuscript/manuscript_pve_editor.Rmd
editor response finished and sent to Remko

```{r global data/sourcing, echo=FALSE, message=FALSE, warning=FALSE}
# Set Flag to include figures in text or not.
includeFigs <- FALSE
```

@@

\
*P* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *g*~1~ and *R*~dark~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above). The *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. *R*~dark~ was was estimated using Eq. (2) from @atkin2005response, which was parameterized with a mean reference empircal measurement for each treatment, a Q~10~ value of 1.86, a reference air temperature for each treatment and air temperatures at each time step.

>>>>

```{r global data/sourcing, echo=FALSE, message=FALSE, warning=FALSE}
# Set Flag to include figures in text or not.
includeFigs <- TRUE
```

@@

\
*P* was predicted by using a coupled photosynthesis - stomatal conductance model [@farquhar1980biochemical;@medlyn2011reconciling] with the 'plantecophys' package [@duursma2015] in R with the mean photosynthetic parameters (*J*~max~, *V*~cmax~, *g*~1~ and *R*~dark~) for each treatment and meteorological data from an onsite weather station. Values of *J*~max~ and *V*~cmax~ are mean estimates from AC~i~ curves over two measurement periods (explained above). The *g*~1~ parameter was generated by fitting the optimal stomatal conductance model from [@medlyn2011reconciling] to observed *g*~s~ values. *R*~dark~ was was estimated using Eq. (2) from @atkin2005response, which was parameterized with a mean reference empirical measurement for each treatment, a Q~10~ value of 1.86, a reference air temperature for each treatment and air temperatures at each time step.
